---
title: "Download White House Visitor Logs for 2024"
author: "Earl F Glynn<br><small>watchdoglab.substack.com/</small>"
date: "<small>`r Sys.Date()`</small>"
output:
  html_document:
    code_download: true
    theme: cerulean
    toc: yes
    toc_depth:  3
    toc_float:
      collapsed:  yes
      smooth_scroll: yes
    number_sections: yes
    code_folding:  show             
---

```{r setup, echo = FALSE}
# http://biostat.mc.vanderbilt.edu/wiki/Main/KnitrHtmlTemplate
require(Hmisc)    # provides knitrSet and other functions
knitrSet(lang = 'markdown',   # If using blogdown: knitrSet(lang='blogdown')
         fig.align = 'left',
         w = 6.5,
         h = 4.5,
         cache = FALSE)
```

`r hidingTOC(buttonLabel = "Outline")`

```{r startYourEngines, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  comment = NA)

time.1 <- Sys.time()
```

[Biden-â Harris Administration Reinstates Visitor Log Policy 2021-05-07](https://www.whitehouse.gov/briefing-room/disclosures/2021/05/07/biden-harris-administration-reinstates-visitor-log-policy-will-be-first-administration-to-post-records-from-first-full-year-in-office/)

* [Voluntary Disclosure Policy for White House Visitor Logs](https://www.whitehouse.gov/voluntary-disclosure/)

* [Visitor Logs](https://www.whitehouse.gov/disclosures/visitor-logs/)

Links to download `.csv` files are in the section `Visitor Logs by Date`

Visitor Log Records Posting Dates from [Briefing Room](https://www.whitehouse.gov/briefing-room/):

* See file `Briefing-Room.docx`

* [January 2024 Visitor Logs Records Posted](https://www.whitehouse.gov/briefing-room/disclosures/2024/04/30/january-2024-visitor-logs-records-posted/)


* [Visitor Logs](https://www.whitehouse.gov/disclosures/visitor-logs/)

# Setup {.tabset .tabset-fade .tabset-pills}

## {.active}

## Constants

```{r}
EXCEL_LIMIT <- 2^20
```

## Packages

```{r Packages}
library(tidyverse)
```

Display formatting

```{r DisplayFormat}
library(kableExtra)   # kable_styling
```

I/O

```{r IO}
library(readxl)      # read_xlsx
library(readr)       # write_csv
library(writexl)     # write_xlsx
```                                          

```{r}
library(downloader)  # download
library(tools)       # md5sum        
```

Display formatting

```{r Display}
library(kableExtra)  # kable_styling
#library(IRdisplay)   # display_html (needed in Jupyter notebooks)
```         

## Helper functions

```{r Helpers}
Show <- function(data, caption="", bigMark="",
                 height = NULL, width = NULL, ...)
{
  data                                       |>
  kable("html", caption=caption,
        format.args=list(big.mark=bigMark))  |>
  kable_styling(bootstrap_options=c("striped", "bordered", "condensed"),
                position="left",
                full_width=FALSE, ...)       |>
  scroll_box(height = height, width = width) 
}
```                           

# Download Month Files in 2022

* https://www.whitehouse.gov/ | Disclosures

* https://www.whitehouse.gov/disclosures/ | Vistor Logs

* https://www.whitehouse.gov/disclosures/visitor-logs/ 

Bureaucrats can't use a simple pattern ... 

* https://www.whitehouse.gov/wp-content/uploads/2022/05/2022.01_WAVES-ACCESS-RECORDS.csv

* https://www.whitehouse.gov/wp-content/uploads/2022/05/2022.02_WAVES-ACCESS-RECORDS.csv

* https://www.whitehouse.gov/wp-content/uploads/2022/07/2022.03_WAVES-ACCESS-RECORDS-.csv

* https://www.whitehouse.gov/wp-content/uploads/2022/08/2022.04_WAVES-ACCESS-RECORDS.csv

* https://www.whitehouse.gov/wp-content/uploads/2022/08/2022.05-WAVES-ACCESS-RECORDS.csv

* https://www.whitehouse.gov/wp-content/uploads/2022/10/2022.06_WAVES-ACCESS-RECORDS.csv

* https://www.whitehouse.gov/wp-content/uploads/2022/10/2022.07_WAVES-ACCESS-RECORDS.csv

* https://www.whitehouse.gov/wp-content/uploads/2022/11/2022.08_WAVES-ACCESS-RECORDS.csv

* https://www.whitehouse.gov/wp-content/uploads/2022/12/2022.09_WAVES-ACCESS-RECORDS.csv

* https://www.whitehouse.gov/wp-content/uploads/2023/01/2022.10_WAVES-ACCESS-RECORDS.csv

* https://www.whitehouse.gov/wp-content/uploads/2023/03/2022.12_WAVES-ACCESS-RECORDS.csv

# Download Month Files in 2023

* https://www.whitehouse.gov/wp-content/uploads/2023/04/2023.01_WAVES-ACCESS-RECORDS.csv

* https://www.whitehouse.gov/wp-content/uploads/2023/05/2023.02_WAVES-ACCESS-RECORDS.csv

* https://www.whitehouse.gov/wp-content/uploads/2023/06/2023.03_WAVES-ACCESS-RECORDS.csv

* https://www.whitehouse.gov/wp-content/uploads/2023/07/2023.04_WAVES-ACCESS-RECORDS.csv 

* https://www.whitehouse.gov/wp-content/uploads/2023/08/2023.05_WAVES-ACCESS-RECORDS.csv

* https://www.whitehouse.gov/wp-content/uploads/2023/09/2023.06_WAVES-ACCESS-RECORDS.csv

* https://www.whitehouse.gov/wp-content/uploads/2023/10/2023.07_WAVES-ACCESS-RECORDS.csv

* https://www.whitehouse.gov/wp-content/uploads/2023/11/2023.08_WAVES-ACCESS-RECORDS.csv

* https://www.whitehouse.gov/wp-content/uploads/2023/12/2023.09_WAVES-ACCESS-RECORDS.csv

* https://www.whitehouse.gov/wp-content/uploads/2024/01/2023.10_WAVES-ACCESS-RECORDS.csv

* https://www.whitehouse.gov/wp-content/uploads/2024/02/2023.11_WAVES-ACCESS-RECORDS.csv 

* https://www.whitehouse.gov/wp-content/uploads/2024/03/2023.12_WAVES-ACCESS-RECORDS.csv

# Download Month Files in 2024

* https://www.whitehouse.gov/wp-content/uploads/2024/04/2024.01_WAVES-ACCESS-RECORDS.csv

Notes:

* extra "-" at end of filename in March 2022.

* "-" not "_" in filename in May 2022.

```{r DownloadFunction}
DownloadMonth <- function(MonthPath, MonthFilename)
{
  BASE_URL <- "https://www.whitehouse.gov/wp-content/uploads/"
  URL      <- paste0(BASE_URL, MonthPath, MonthFilename)
  download(URL, MonthFilename, mode="wb", quiet=TRUE)  
}
```

```
{r}
DownloadMonth("2022/05/", "2022.01_WAVES-ACCESS-RECORDS.csv")  # Jan 22
DownloadMonth("2022/05/", "2022.02_WAVES-ACCESS-RECORDS.csv")  # Feb 22
DownloadMonth("2022/07/", "2022.03_WAVES-ACCESS-RECORDS-.csv") # Mar 22   extra "-"
DownloadMonth("2022/08/", "2022.04_WAVES-ACCESS-RECORDS.csv")  # Apr 22
DownloadMonth("2022/08/", "2022.05-WAVES-ACCESS-RECORDS.csv")  # May 22   "-" not "_" in name
DownloadMonth("2022/10/", "2022.06_WAVES-ACCESS-RECORDS.csv")  # Jun 22
DownloadMonth("2022/10/", "2022.07_WAVES-ACCESS-RECORDS.csv")  # Jul 22
DownloadMonth("2022/11/", "2022.08_WAVES-ACCESS-RECORDS.csv")  # Aug 22
DownloadMonth("2022/12/", "2022.09_WAVES-ACCESS-RECORDS.csv")  # Sep 22
DownloadMonth("2023/01/", "2022.10_WAVES-ACCESS-RECORDS.csv")  # Oct 22
DownloadMonth("2023/02/", "2022.11_WAVES-ACCESS-RECORDS.csv")  # Nov 22
DownloadMonth("2023/03/", "2022.12_WAVES-ACCESS-RECORDS.csv")  # Dec 22
```

```
DownloadMonth("2023/04/", "2023.01_WAVES-ACCESS-RECORDS.csv") # Jan 23
DownloadMonth("2023/05/", "2023.02_WAVES-ACCESS-RECORDS.csv") # Feb 23
DownloadMonth("2023/06/", "2023.03_WAVES-ACCESS-RECORDS.csv") # Mar 23
DownloadMonth("2023/07/", "2023.04_WAVES-ACCESS-RECORDS.csv") # Apr 23
DownloadMonth("2023/08/", "2023.05_WAVES-ACCESS-RECORDS.csv") # May 23
DownloadMonth("2023/09/", "2023.06_WAVES-ACCESS-RECORDS.csv") # June 23
DownloadMonth("2023/10/", "2023.07_WAVES-ACCESS-RECORDS.csv") # July 23
DownloadMonth("2023/11/", "2023.08_WAVES-ACCESS-RECORDS.csv") # Aug 23
DownloadMonth("2023/12/", "2023.09_WAVES-ACCESS-RECORDS.csv") # Sept 23
DownloadMonth("2024/01/", "2023.10_WAVES-ACCESS-RECORDS.csv") # Oct 23
DownloadMonth("2024/02/", "2023.11_WAVES-ACCESS-RECORDS.csv") # Nov 23
DownloadMonth("2024/03/", "2023.12_WAVES-ACCESS-RECORDS.csv") # Dec 23
```

```
DownloadMonth("2024/04/", "2024.01_WAVES-ACCESS-RECORDS.csv") # Jan 24
```

# Download Month Files in 2023

```{r}
DownloadMonth("2024/06/", "2024.02_WAVES-ACCESS-RECORDS.csv") # Feb 24
```


Modify `filenamePattern` to accommodate March '22 with the extra dash

```{r}
filenamePattern <- "^2024\\.[0-9]{2}.WAVES-ACCESS-RECORDS"    ######
filenames       <- dir(pattern = filenamePattern)

metaInfoFilenames <- 
  file.info(filenames)           |>      # metadata about file(s) 
  rownames_to_column("Filename") |>
  select(-isdir, -mode, -exe)
```

```{r}
metaInfoFilenames$md5sum <- md5sum(filenames) 
```

```{r}
metaInfoFilenames |> Show()
```

```{r}
stamp <- format(time.1, "%Y-%m-%d")
write_xlsx(metaInfoFilenames, 
           paste0("White-House-Visitor-Logs-2023-", stamp, ".xlsx"))
```


# Epilog {.tabset .tabset-fade .tabset-pills}

## {.active}

## Session Info

```{r devtoolsSessionInfo}
devtools::session_info()
```

</div>

```{r epilogDeltaTime, echo=FALSE}
time.2 <- Sys.time()
processingTime <- paste("Processing time:", sprintf("%.1f",
                        as.numeric(difftime(time.2,
                                            time.1, units="secs"))), "secs\n")
```

`r processingTime`
`r format(time.2, "%Y-%m-%d %H%M")`

