---
title: "White House Visitors - Dec 2021 `Gifts from the Heart` Visitors"
author: "Earl F Glynn<br>"
date: "<small>`r Sys.Date()`</small>"
output:
  html_document:
    code_download: true
    theme: cerulean
    toc: yes
    toc_depth:  3
    toc_float:
      collapsed:  yes
      smooth_scroll: yes
    number_sections: yes
    code_folding:  show     
---

```{r setup, echo = FALSE}
# http://biostat.mc.vanderbilt.edu/wiki/Main/KnitrHtmlTemplate
require(Hmisc)    # provides knitrSet and other functions
knitrSet(lang = 'markdown',   # If using blogdown: knitrSet(lang='blogdown')
         fig.align = 'left',
         w = 6.5,
         h = 4.5,
         cache = FALSE)
```

`r hidingTOC(buttonLabel = "Outline")`

```{r startYourEngines, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  comment = NA)

time.1 <- Sys.time()
```

# Setup {.tabset .tabset-fade .tabset-pills}

## {.active}

## Constants

```{r}
DATA_RELEASE  <- "2023-01-30"          #####
```

[October 2022 Visitor Log Records Posted, Jan. 30, 2023](https://www.whitehouse.gov/briefing-room/disclosures/2023/01/30/october-2022-visitor-log-records-posted/)

```{r}
SOURCE <-  paste0("Source: www.whitehouse.gov/briefing-room/disclosures/2023/01/30/\n",
                  "october-2022-visitor-log-records-posted/")

stamp   <- paste0("watchdoglab.substack.com\n", format(time.1, "%Y-%m-%d"))
```       

```{r}
EXCEL_LIMIT <- 2^20
```

## Packages

```{r}
library(tidyverse)
library(lubridate)   # parse_date_time, mdy_hm

library(downloader)  # download
library(tools)       # md5sum        
```

I/O

```{r IO}
library(readxl)      # read_xlsx
library(readr)       # write_csv
library(writexl)     # write_xlsx
```

Display formatting

```{r Display}
library(kableExtra)  # kable_styling
#library(IRdisplay)   # display_html (needed in Jupyter notebooks)
```         
Graphics

```{r Graphics}
library(RColorBrewer)
library(scales)      # comma, comma_format
```

## ggplot theme

[ggplot Themes Gallery](https://www.datanovia.com/en/blog/ggplot-themes-gallery/)

```{r ggplotTheme}
theme_set(theme_minimal() +

            theme(
                   axis.text             = element_text(size =  9),
                   axis.title            = element_text(size = 11),

                   plot.caption          = element_text(size =  9,
                                                        hjust = c(0.0,1.0)),
                   plot.caption.position = "plot",  
  
                   plot.title.position   = "plot",
                   plot.title            = element_text(size = 14),
                   plot.subtitle         = element_text(size = 12),
                   
                   strip.background      = element_rect(fill = "aliceblue"),
                   strip.text            = element_text(size = 14),             
            
                   panel.grid.minor.x    = element_blank(),
                   panel.grid.minor.y    = element_blank()   
                 )
         )

COLOR_BAR     <- "skyblue"
COLOR_OUTLINE <- "grey80"
```                                      

## Helper functions

### Show

```{r Helpers}
Show <- function(data, caption="", bigMark="",
                 height = NULL, width = NULL, ...)
{
  data                                       |>
  kable("html", caption=caption,
        format.args=list(big.mark=bigMark))  |>
  kable_styling(bootstrap_options=c("striped", "bordered", "condensed"),
                position="left",
                full_width=FALSE, ...)       |>
  scroll_box(height = height, width = width) 
}
```       

# Read White House Visitor Data

Combined from all Biden releases in script
`1-WhiteHouse-Visitors-Processing.html`

```{r}
wh <- 
  read_csv(paste0("White-House-Visitor-Log-Data-",
                  DATA_RELEASE,
                  ".csv"),
           guess_max = EXCEL_LIMIT,
           show_col_types = FALSE)   |>
  select(-`Terminal Suffix`)         |>       # identical to `Last Updated By`
  mutate(GROUP = factor(GROUP,                # Define order for groups
                        levels = c("POTUS", "VPOTUS", "FLOTUS",
                                   "Staff", "Tourist")))
    
```

```{r}
glimpse(wh)
```

# Tourist Visitors for Dec 2021

```{r}
dec2021 <- 
  wh |>
  filter(APPT_START_DATE  > "2021-11-30",
         APPT_START_DATE  < "2022-01-01",
         GROUP           == "Tourist")

dim(dec2021)
```

```{r}
dec2021 |> count(VISITEE) |> Show()
```

```{r}
dec2021 |> count(LOCATION) |> Show()
```

```{r}
dec2021 |> count(CALLER) |> Show()
```

```{r}
write_xlsx(dec2021,
           paste0("White-House-Tourist-Visitors-for-Dec-2021-",
                  DATA_RELEASE, ".xlsx"))
```      

# Event Summary

```{r}
decemberTouristEvents <-
  dec2021 |>
  group_by(APPT_START_DATE, LOCATION, Event, `Total People`) |>
  summarize(ActualCount = n(),
            CancelCount = sum(!is.na(APPT_CANCEL_DATE)),
            .groups = "drop")

dim(decemberTouristEvents)
```

```{r}
decemberTouristEvents |> head(3) |> Show()
```

Size of Groups

```{r}
fivenum(decemberTouristEvents$ActualCount)
```

Over half of the events were for one or two persons.

```{r}
table(decemberTouristEvents$ActualCount)
```

Canceled Events

```{r}
decemberTouristEvents |> filter(CancelCount > 0) |> Show()
```

```{r}
write_xlsx(decemberTouristEvents,
           paste0("White-House-Tourist-Events-for-Dec-2021-",
                  DATA_RELEASE, ".xlsx"))
```      
      
# Summarize Tourist Counts by Date

```{r}
byDate <- 
  dec2021                                        |>
  mutate(APPT_START_DATE = ymd(APPT_START_DATE)) |>
  group_by(APPT_START_DATE)                      |>
  summarize(Visitors = n(),
            Events   = n_distinct(Event))
  
nrow(byDate)
```


```{r}
byDate |> Show(bigMark = ",", height = "300px")
```

```{r}
c(sum(byDate$Visitors), sum(byDate$Events))
```

```{r}
set3 <- brewer.pal(12, "Set3")
VISITOR_COLOR <- set3[9]
```


```{r DecHolidaysbyDate}
ggplot(byDate, aes(x = APPT_START_DATE, y = Visitors))      +
  geom_bar(stat = "identity", fill = VISITOR_COLOR)         +
   scale_x_date(
                date_breaks = "1 day",
                labels = date_format("%a\n%b %d")
              )                                             +  
  geom_text(aes(x = APPT_START_DATE, y = Visitors, 
                label = comma(Visitors)),
                vjust = -0.5, hjust = 0.5, 
                size = 3, angle = 0)                        +
  labs(
        title    = "Visitors for 'Gifts from the Heart' Holiday Tour at White House Dec. 2021",
        subtitle = paste0("By Date [",
                          comma(sum(byDate$Visitors)), " total]"),
        x = "Date",
        y = "Visitors",
        caption = c(SOURCE, stamp)
      )                                                     +
  theme(panel.grid.major.x = element_blank())
```

# Summarize Tourist Counts by Time of Day

APPT_START_DATE is missing times, but Time of Arrival (TOA) is a good proxy.

```{r}
hourCountsTotal <- 
  dec2021                |> 
  summarize(n            = n(),
            hoursMissing = sum(is.na(TOAConverted))) 

hourCountsTotal |> Show()
```

```{r}
byHour <- 
  dec2021                           |>
  mutate(Hour = hour(TOAConverted)) |>
  filter(!is.na(Hour))              |>
  group_by(Hour)                    |>
  summarize(Visitors = n(),
            Events   = n_distinct(Event))
  
nrow(byHour)
```

```{r}
byHour |> Show()
```

```{r DecHolidaysbyHour, fig.width = 7, fig.height = 7}
 ggplot(byHour, aes(x = Hour, y = Visitors))              +
  coord_polar(theta = "x",   direction = 1)               +
  geom_bar(stat = "identity", fill = VISITOR_COLOR, 
           color = COLOR_OUTLINE)                         +
  scale_x_continuous(breaks = seq(0, 24), 
                     limits = c(0, 24),
                     labels = seq(0, 24))                 +
  scale_y_continuous(breaks = 250 * 0:4,
                     limits = c(0, 1000))                 +
  labs(title    = "Visitors for 'Gifts from the Heart' Holiday Tour at White House Dec. 2021",
       subtitle = paste0("By Hour of Day [based on arrival times of ",
                        comma(hourCountsTotal$n - hourCountsTotal$hoursMissing),
                        " of ", comma(hourCountsTotal$n), " visitors]"),
       x = "Hour of Day", 
       y = "Visitors", 
       caption = c(SOURCE, stamp))                        +
  theme(axis.text  = element_text(size = 10),
        axis.title = element_text(size = 11))
```

# Epilog {.tabset .tabset-fade .tabset-pills}

## {.active}

## Session Info

```{r devtoolsSessionInfo}
devtools::session_info()
```

</div>

```{r epilogDeltaTime, echo=FALSE}
time.2 <- Sys.time()
processingTime <- paste("Processing time:", sprintf("%.1f",
                        as.numeric(difftime(time.2,
                                            time.1, units="secs"))), "secs\n")
```

`r processingTime`
`r format(time.2, "%Y-%m-%d %H%M")`
