---
title: "White House Visitors - Explorations - By Day for Latest Month"
author: "Earl F Glynn<br>"
date: "<small>`r Sys.Date()`</small>"
output:
  html_document:
    code_download: true
    theme: cerulean
    toc: yes
    toc_depth:  3
    toc_float:
      collapsed:  yes
      smooth_scroll: yes
    number_sections: yes
    code_folding:  show     
---

Also see `Counts` directory created by script `1-WhiteHouse-Visitors-Processing.Rmd` with frequency counts of values of fields.

```{r setup, echo = FALSE}
# http://biostat.mc.vanderbilt.edu/wiki/Main/KnitrHtmlTemplate
require(Hmisc)    # provides knitrSet and other functions
knitrSet(lang = 'markdown',   # If using blogdown: knitrSet(lang='blogdown')
         fig.align = 'left',
         w = 6.5,
         h = 4.5,
         cache = FALSE)
```

`r hidingTOC(buttonLabel = "Outline")`

```{r startYourEngines, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  comment = NA)

time.1 <- Sys.time()
```

# Setup {.tabset .tabset-fade .tabset-pills}

## {.active}

## Constants

```{r}
DATA_RELEASE  <- "2023-04-30"                                                              ##### (1/4)
```

[January 2023 Visitor Log Records Posted, April 30, 2023](https://www.whitehouse.gov/briefing-room/disclosures/2023/04/30/january-2023-visitor-log-records-posted/)


```{r}
SOURCE <-  paste0("Source: www.whitehouse.gov/briefing-room/disclosures/2023/04/30/\n",    #####  (2/4)
                  "january-2023-visitor-log-records-posted/")                              #####  (3/4)

stamp   <- paste0("watchdoglab.substack.com\n", format(time.1, "%Y-%m-%d"))
```       


```{r}
EXCEL_LIMIT <- 2^20
```

## Packages

```{r}
library(tidyverse)
library(lubridate)   # parse_date_time, mdy_hm

library(downloader)  # download
library(tools)       # md5sum        
```

I/O

```{r IO}
library(readxl)      # read_xlsx
library(readr)       # write_csv
library(writexl)     # write_xlsx
```

Display formatting

```{r Display}
library(kableExtra)  # kable_styling
#library(IRdisplay)   # display_html (needed in Jupyter notebooks)
```         
Graphics

```{r Graphics}
library(RColorBrewer)
library(scales)      # comma, comma_format
```

## ggplot theme

[ggplot Themes Gallery](https://www.datanovia.com/en/blog/ggplot-themes-gallery/)

```{r ggplotTheme}
theme_set(theme_minimal() +

            theme(
                   axis.text             = element_text(size = 10),
                   axis.title            = element_text(size = 12),

                   plot.caption          = element_text(size =  10,
                                                        hjust = c(0.0,1.0)),
                   plot.caption.position = "plot",  
  
                   plot.title.position   = "plot",
                   plot.title            = element_text(size = 14),
                   plot.subtitle         = element_text(size = 12),
                   
                   strip.background      = element_rect(fill = "aliceblue"),
                   strip.text            = element_text(size = 14),             
            
                   panel.grid.minor.x    = element_blank(),
                   panel.grid.minor.y    = element_blank()   
                 )
         )

COLOR_BAR     <- "skyblue"
COLOR_OUTLINE <- "grey80"
```                                      

## Helper functions

### Show

```{r Helpers}
Show <- function(data, caption="", bigMark="",
                 height = NULL, width = NULL, ...)
{
  data                                       |>
  kable("html", caption=caption,
        format.args=list(big.mark=bigMark))  |>
  kable_styling(bootstrap_options=c("striped", "bordered", "condensed"),
                position="left",
                full_width=FALSE, ...)       |>
  scroll_box(height = height, width = width) 
}
```       

# Read White House Visitor Data

## All Data

Combined from all reports

```{r}
wh <- 
  read_csv(paste0("White-House-Visitor-Log-Data-",
                  DATA_RELEASE,
                  ".csv"),
           guess_max = EXCEL_LIMIT,
           show_col_types = FALSE)   |>
  select(-`Terminal Suffix`)         |>       # identical to `Last Updated By`
  mutate(GROUP = factor(GROUP,                # Define order for groups
                        levels = c("POTUS", "VPOTUS", "FLOTUS",
                                   "Staff", "Tourist")))
    
dim(wh)
```
## Counts by Month

```{r}
monthCounts <-
  wh |> 
    filter(!is.na(RELEASEDATE)) |>  # ignore two records with missing value
    group_by(RELEASEDATE) %>%
    count()

monthCounts |> Show(height = "200px")
```

Normally will want the most recent month, but the filter here could be changed to pick any month.

```{r}
targetRelease <- monthCounts |> select(RELEASEDATE) |> tail(1) |> pull()        ##### 4/4
targetRelease 
```
For filenames

```{r}
targetMonth <-  targetRelease %m-% months(3) |> str_sub(1, 7)                 
targetMonth
```

```{r}
targetMonthName <- format(targetRelease %m-% months(3), "%B %Y")
targetMonthName
```

```{r}
wh <- 
  wh  |>
  filter(RELEASEDATE == targetRelease)

dim(wh)
```

# Define Events

Unique events defined by `Event`

```{r}
events <- 
  wh  |>
  select(GROUP, APPT_START_DATE, 
         Event, LOCATION, `Total People`)                    |>
  unique()                                                   |>
  mutate(              # YYYY-MM-DD yy:mm:ss
          Year       = substr(APPT_START_DATE,  1,  4) |>
                       as.numeric(),
          Month      = substr(APPT_START_DATE,  6,  7) |>
                       as.numeric(),                   
          Day        = substr(APPT_START_DATE,  9, 10) |>
                       as.numeric(),   
          Hour       = substr(APPT_START_DATE, 12, 13) |>
                       as.numeric()
        ) 

dim(events)
```

```{r}
length(unique(wh$Event))
```

```{r}
events |> tail(3) |> Show()
```

```{r}
write_xlsx(events, 
           paste0("Target-", targetMonth, "-",
                  "White-House-Events-", 
                  DATA_RELEASE, ".xlsx"))
```

## Events by Group

```{r}
eventCounts <- table(events$GROUP, useNA = "ifany")
eventCounts
```
```{r}
eventCountsByGroup <- 
  events                      |>
  group_by(`Total People`, 
           GROUP)             |>
  summarize(n = n(),
            .groups = "drop") |> 
  spread(GROUP, n, fill = 0)  |>
  mutate(Events = rowSums(across(POTUS:Tourist)))

dim(eventCountsByGroup)
```

```{r}
eventCountsByGroup |>
  Show(bigMark = ",", height = "300px")
```

```{r}
write_xlsx(eventCountsByGroup, 
           paste0("Target-", targetMonth, "-",
                  "White-House-Events-TotalPeople-by-Group-", 
                  DATA_RELEASE, ".xlsx"))
```

## Events By Day

```{r}
eventCountsByDay <- 
  events                      |>
  group_by(Day, GROUP)        |>
  summarize(n = n(),
            .groups = "drop") |> 
  spread(GROUP, n, fill = 0)  |>
  mutate(Events = rowSums(across(POTUS:Tourist)))

dim(eventCountsByDay)
```

```{r}
if (!("FLOTUS" %in% colnames(eventCountsByDay)))
{
  eventCountsByDay <- 
    eventCountsByDay   |>
    mutate(FLOTUS = 0) |>
    relocate(FLOTUS, .after = VPOTUS)
}
```

```{r}
eventCountsByDay <- 
  eventCountsByDay |>
  mutate( 
          PVF      = POTUS + VPOTUS + FLOTUS
        )               
  

dim(eventCountsByDay)
```

```{r}
eventCountsByDay |>
  Show(bigMark = ",", height = "300px")
```

```{r}
write_xlsx(eventCountsByDay,
           paste0("Target-", targetMonth, "-",
                  "White-House-Events-by-Day-Table-",
                  DATA_RELEASE, ".xlsx"))
```


```{r}
longFormat <- 
  eventCountsByDay        |>
  pivot_longer(
                cols      = c("POTUS",  "VPOTUS", "FLOTUS", 
                              "Staff",  "Tourist",
                              "Events", "PVF"),
                names_to  = "Group",
                values_to = "Events"
               )

dim(longFormat)
```
### PVL by Day

Use default blue color for POTUS, but two different sets of qualitative pairs for plots below.

```{r}
POTUS_COLOR <- "#619CFF"
set2 <- brewer.pal( 8, "Set2")  # Brewer qualitative palettes
set3 <- brewer.pal(12, "Set3")
```

```{r EventsBidenPotusVPotusFlotus, fig.height = 4, fig.width = 8}
longFormat                                                               |>
  filter(Group %in% c("POTUS", "VPOTUS", "FLOTUS"))                      |>
  mutate(Group = factor(Group, levels = c("FLOTUS", "VPOTUS", "POTUS"))) |>
  ggplot(aes(x = Day, y = Events, fill = Group))        +
  geom_bar(position = "stack", stat = "identity")       +
  scale_x_continuous(breaks = 1:nrow(eventCountsByDay)) +
  scale_y_continuous(
                      labels = scales::comma
                    )                                   +
                             # FLOTUS    VPOTUS    POTUS
  scale_fill_manual(values = c(set2[5],  set2[6],  POTUS_COLOR))  +

  labs(
        title    = paste0("White House Events During Biden Administration in ", targetMonthName),  
        subtitle = paste0("Event Counts by Day for POTUS, VPOTUS, FLOTUS "),
        x = "Day of Month", 
        y = "Events",
        caption = c(SOURCE, stamp)
      )                                                 +                                   
  theme(panel.grid.major.x    = element_blank(),
        legend.position = "right")
```

```{r EventsBidenPotusStaffTourist, fig.height = 4, fig.width = 8}
longFormat                                                             |>
  filter(Group %in% c("PVF", "Staff", "Tourist"))                      |>
  mutate(Group = factor(Group, levels = c("Tourist", "Staff", "PVF"))) |>
  ggplot(aes(x = Day, y = Events, fill = Group))        +
  geom_bar(position = "stack", stat = "identity")       +
  
  scale_y_continuous(labels = scales::comma)            +
  scale_x_continuous(breaks = 1:nrow(eventCountsByDay)) +
 
   scale_fill_manual(values=c(set3[9],  set3[10], POTUS_COLOR),
                    labels = c("Tourist", "Staff", "POTUS,\nVPOTUS,\nFLOTUS"))  + 
  
  labs(
        title    = paste0("White House Events During Biden Administration in ", targetMonthName),  
        subtitle = "Event Counts by Day for POTUS/VPOTUS/FLOTUS, Staff, Tourists",
        x = "Day of Month",
        y = "Events",
        caption = c(SOURCE, stamp)
      )                                             +                                   
  theme(panel.grid.major.x    = element_blank(),
        legend.position = "right")
```

### Staff by Day



# Total People per Event by Number of Events

## All Events

Use outline color black to show the 1 levels, and overlaps.

Don't bother with histograms here, since log-log scale already difficult to interpret.

```{r TargetTotalPeopleAllEvents}
ggplot(eventCountsByGroup, aes(x = `Total People`, y= Events))  + 
  geom_bar(stat = "identity", 
           fill = COLOR_BAR,  color = "black", width = 0.02)    +        
  scale_x_log10(breaks = c(1:5, 10, 20, 30, 50,
                           100, 200, 500, 1000, 2000, 5000),
                labels = scales::comma)                         +
  scale_y_log10(breaks = c(1, 5, 10, 50, 100, 500,
                           1000, 5000, 10000, 15000, 20000),
                labels = scales::comma)                         +
  labs(
        title    = "Total People per White House Event",
        subtitle = paste0("All Visitors, ", comma(nrow(events)), 
                          " Events in ", targetMonthName),
        x = "Total People Per Event",
        y = "Events",
        caption = c(SOURCE, stamp)
      )                                                  

```

## POTUS Events

Not sure why, but x-axis scaling is not very dynamic unless filtered to only GROUP of interest.

```{r TargetTotalPeoplePOTUSEvents}
eventCountsByGroup               |> 
  select(`Total People`, POTUS)  |>
  filter(POTUS > 0)              |>
ggplot(aes(x = `Total People`, y= POTUS))                             + 
  geom_bar(stat = "identity", 
           fill = COLOR_BAR,  color = "black", width = 0.025)         +        
  scale_x_log10(breaks = c(1:5, 10, 20, 30, 50,
                           100, 200, 500, 1000, 2000, 5000),
                labels = scales::comma)                               +
  scale_y_log10(breaks = c(1, 5, 10, 50, 100, 500,
                           1000, 5000, 10000, 15000, 20000),
                labels = scales::comma)                               +
  labs(
        title    = "Total People per White House Event - POTUS",
        subtitle = paste0("POTUS (President of the United States) Visitors, ",
                          comma(sum(eventCountsByGroup$POTUS)), 
                          " Events in ", targetMonthName),
        x = "Total People Per Event",
        y = "Events",
        caption = c(SOURCE, stamp)
      )                                        
```

## VPOTUS Events

```{r TargetTotalPeopleVPOTUSEvents}
eventCountsByGroup               |> 
  select(`Total People`, VPOTUS) |>
  filter(VPOTUS > 0)             |>
ggplot(aes(x = `Total People`, y= VPOTUS))                            + 
  geom_bar(stat = "identity", 
           fill = COLOR_BAR, color = "black", width = 0.05)           +        
  scale_x_log10(breaks = c(1:5, 10, 20, 30, 50,
                           100, 150, 300),
                labels = scales::comma)                               +
  scale_y_log10(#breaks = c(1, 5, 10, 20, 30, 40),
                labels = scales::comma)                               +
  labs(
        title    = "Total People per White House Event - VPOTUS",
        subtitle = paste0("VPOTUS (Vice-President of the United States) Visitors, ",
                          comma(sum(eventCountsByGroup$VPOTUS)), 
                          " Events in ", targetMonthName),
        x = "Total People Per Event",
        y = "Events",
        caption = c(SOURCE, stamp)
      )                                        
```

## FLOTUS Events

```{r TargetTotalPeopleFLOTUSEvents}
if (eventCounts["FLOTUS"] > 0)
{
  eventCountsByGroup                |> 
    select(`Total People`, FLOTUS)  |>
    filter(FLOTUS > 0)              |>
  ggplot(aes(x = `Total People`, y= FLOTUS))                            + 
    geom_bar(stat = "identity", 
             fill = COLOR_BAR, color = "black", width = 0.05)           +        
    scale_x_log10(breaks = c(1:5, 10, 20, 30, 50,
                             100, 150),
                  labels = scales::comma)                               +
    scale_y_log10(# breaks = c(1, 5, 10, 20, 30, 40),
                  labels = scales::comma)                               +
    labs(
          title    = "Total People per White House Event - FLOTUS",
           subtitle = paste0("FLOTUS (First Lady of the United States) Visitors, ",
                            comma(sum(eventCountsByGroup$FLOTUS)), 
                             " Events in ", targetMonthName),
          x = "Total People Per Event",
          y = "Events",
          caption = c(SOURCE, stamp)
        ) 
}
```

## Staff Events

```{r TargetTotalPeopleStaffEvents}
eventCountsByGroup               |> 
  select(`Total People`, Staff)  |>
  filter(Staff > 0)              |>
ggplot(aes(x = `Total People`, y= Staff))                             + 
  geom_bar(stat = "identity", 
           fill = COLOR_BAR,  color = "black", width = 0.025)         +        
  scale_x_log10(breaks = c(1:5, 10, 20, 30, 50,
                           100, 250, 500, 750, 1250),
                labels = scales::comma)                               +
  scale_y_log10(breaks = c(1, 5, 10, 50, 100, 500,
                           1000, 5000, 10000, 15000, 20000),
                labels = scales::comma)                               +
  labs(
        title    = "Total People per White House Event - Staff",
        subtitle = paste0("White House Staff Visitors, ",
                          comma(sum(eventCountsByGroup$Staff)), 
                          " Events in ", targetMonthName),
        x = "Total People Per Event",
        y = "Events",
        caption = c(SOURCE, stamp)
      )                                        
```

```{r TargetTotalPeopleTouristEvents}
eventCountsByGroup                 |> 
  select(`Total People`, Tourist)  |>
  filter(Tourist > 0)              |>
ggplot(aes(x = `Total People`, y= Tourist))                           + 
  geom_bar(stat = "identity", 
           fill = COLOR_BAR,  color = "black", width = 0.025)         +        
  scale_x_log10(breaks = c(1:5, 10, 20, 30, 50, 75, 100, 125),
                labels = scales::comma)                               +
  scale_y_log10(breaks = c(1, 5, 10, 50, 100, 500,
                           1000, 2000, 4000, 5000, 7000),
                labels = scales::comma)                               +
  labs(
        title    = "Total People per White House Event - Tourists",
        subtitle = paste0("White House Tourist Visitors, ",
                          comma(sum(eventCountsByGroup$Tourist)),
                          " Events in ", targetMonthName),
        x = "Total People Per Event",
        y = "Events",
        caption = c(SOURCE, stamp)
      )                                        
```

# Event Start Date Hour

Next time:  use function for the Hour plots

## Overall

Most of the Hour = 0 values are caused by missing time fields with date-only values.

```{r}
hourCounts <- 
  events         |>
  group_by(Hour) |>
  count()

nrow(hourCounts)
```
```{r}
sum(hourCounts$n)
```

Would like to use `coord_polar`for 24-hour clock here but it is fickle or has bugs.

Oddly, must expand left x-limit or first bar is clipped when present -- this as also a problem with `coord_polar`

```{r}
hourCountsLessMissing <- 
  hourCounts  |>
  filter(Hour != 0)

sum(hourCountsLessMissing$n)
```

```{r TargetEventHourAll}    
ggplot(hourCountsLessMissing)          +
  geom_bar(aes(x = Hour, y = n),
           stat = "identity", 
           fill = COLOR_BAR,  color = "black")               + 
  
  scale_x_continuous(limits = c(0, 23),
                     breaks = seq(0, 23))                    +
  scale_y_continuous(labels = scales::comma)                 +
  
  geom_text(aes(x = Hour, y = n, label = comma(n)),
            vjust = -0.5, hjust = 0.5, size = 2, angle = 0) +
  
  labs(title = "White House Event Start Time Hour (24-hour clock)",
       subtitle = paste0(comma(sum(hourCountsLessMissing$n)), " events [",
                         comma(sum(hourCounts$n) -
                               sum(hourCountsLessMissing$n)),
                         " events missing time value] ",
                         "in ", targetMonthName),
       x = "Hour of Day", 
       y = "Events", 
       caption = c(SOURCE, stamp))                           +
  theme(panel.grid.major.x    = element_blank())
```

Unusual Start Times

```{r}
unusualStart <- 
  events                        |>
  filter(Hour < 6 | Hour > 22)  |>
  select(Event)                 |>
  inner_join(wh, by = "Event")  |>
  arrange(Event, APPT_START_DATE)

dim(unusualStart)
```

```{r}
write_xlsx(unusualStart,
           paste0("Unusual-Event-Start-", targetMonth, "-",
                  DATA_RELEASE, ".xlsx"))
```

## By Group

```{r}
hourCountsByGroup <- 
  events                                         |>
  group_by(GROUP)                                |>
  count()
  
hourCountsByGroup |> Show(height = "200px")
```

```{r}
hourCountsByGroupYear <- 
  events                                            |>
  mutate(Year = substr(APPT_START_DATE,  1,  4)  |>
                as.numeric(),
         Hour = substr(APPT_START_DATE, 12, 13)  |>
                as.numeric())                        |>
  group_by(GROUP, Year)                             |> 
  count(Hour)
  
hourCountsByGroupYear |> Show(height = "200px")
```

```{r}
write_xlsx(hourCountsByGroupYear, 
           paste0("Target-", targetMonth, "-",
                  "White-House-Event-Start-Hour-by-Group-",
                  DATA_RELEASE, ".xlsx"))
```

### POTUS

```{r}
hourCounts <-
  events                   |>
  filter(GROUP == "POTUS") |>
  group_by(Hour)           |>
  count()

nrow(hourCounts)
```

```{r}
sum(hourCounts$n)
```

```{r}
hourCountsLessMissing <-
  hourCounts  |>
  filter(Hour != 0)

sum(hourCountsLessMissing$n)
```                                     

```{r TargetEventHourPotus}
ggplot(hourCountsLessMissing)          +
  geom_bar(aes(x = Hour, y = n),
           stat = "identity",
           fill = COLOR_BAR,  color = "black")               +
  
   scale_x_continuous(limits = c(0, 23),
                     breaks = seq(0, 23))                    +
  scale_y_continuous(labels = scales::comma)                 +

  geom_text(aes(x = Hour, y = n, label = comma(n)),
            vjust = -0.5, hjust = 0.5, size = 2, angle = 0) +

  labs(title = "White House Event Start Time Hour - POTUS",
       subtitle = paste0(comma(sum(hourCountsLessMissing$n)), " POTUS events [",
                         comma(sum(hourCounts$n) -
                               sum(hourCountsLessMissing$n)),
                         " events missing time value] ",
                         "in Month ", targetMonthName),
       x = "Hour of Day",
       y = "Events",
       caption = c(SOURCE, stamp))                           +
  theme(panel.grid.major.x    = element_blank())
```                                                                 

### VPOTUS

```{r}
hourCounts <-
  events                    |>
  filter(GROUP == "VPOTUS") |>
  group_by(Hour)            |>
  count()

nrow(hourCounts)
```

```{r}
sum(hourCounts$n)
```

```{r}
hourCountsLessMissing <-
  hourCounts  |>
  filter(Hour != 0)

sum(hourCountsLessMissing$n)
```

```{r TargetEventHourVPotus}
ggplot(hourCountsLessMissing)          +
  geom_bar(aes(x = Hour, y = n),
           stat = "identity",
           fill = COLOR_BAR,  color = "black")               +
  
  scale_x_continuous(limits = c(0, 23),
                     breaks = seq(0, 23))                    +
  scale_y_continuous(labels = scales::comma)                 +

  geom_text(aes(x = Hour, y = n, label = comma(n)),
            vjust = -0.5, hjust = 0.5, size = 2, angle = 0) +

  labs(title = "White House Event Start Time Hour - VPOTUS",
       subtitle = paste0(comma(sum(hourCountsLessMissing$n)), " VPOTUS events [",
                         comma(sum(hourCounts$n) -
                               sum(hourCountsLessMissing$n)),
                         " events missing time value] ",
                         "in ", targetMonthName),
       x = "Hour of Day",
       y = "Events",
       caption = c(SOURCE, stamp))                           +
  theme(panel.grid.major.x    = element_blank())
```                                                                               

### FLOTUS

```{r}
hourCounts <-
  events                    |>
  filter(GROUP == "FLOTUS") |>
  group_by(Hour)            |>
  count()

nrow(hourCounts)
```

```{r}
sum(hourCounts$n)
```

```{r}
hourCountsLessMissing <-
  hourCounts  |>
  filter(Hour != 0)

sum(hourCountsLessMissing$n)
```

```{r TargetEventHourFlotus}
if (nrow(hourCounts) > 0)
{
  ggplot(hourCountsLessMissing)          +
    geom_bar(aes(x = Hour, y = n),
             stat = "identity",
             fill = COLOR_BAR,  color = "black")               +
    
    scale_x_continuous(limits = c(0, 23),
                       breaks = seq(0, 23))                    +
    scale_y_continuous(labels = scales::comma)                 +
  
    geom_text(aes(x = Hour, y = n, label = comma(n)),
              vjust = -0.5, hjust = 0.5, size = 2, angle = 0) +
  
    labs(title = "White House Event Start Time Hour - FLOTUS",
         subtitle = paste0(comma(sum(hourCountsLessMissing$n)), " FLOTUS events [",
                           comma(sum(hourCounts$n) -
                                 sum(hourCountsLessMissing$n)),
                           " events missing time value] ",
                           "in ", targetMonthName),
         x = "Hour of Day",
         y = "Events",
         caption = c(SOURCE, stamp))                           +
    theme(panel.grid.major.x    = element_blank())
}
```                                                               

### Staff

```{r}
hourCounts <-
  events                   |>
  filter(GROUP == "Staff") |>
  group_by(Hour)           |>
  count()

nrow(hourCounts)
```

```{r}
sum(hourCounts$n)
```

```{r}
hourCountsLessMissing <-
  hourCounts  |>
  filter(Hour != 0)

sum(hourCountsLessMissing$n)
```

```{r TargetEventHourStaff}
ggplot(hourCountsLessMissing)          +
  geom_bar(aes(x = Hour, y = n),
           stat = "identity",
           fill = COLOR_BAR,  color = "black")               +
  
  scale_x_continuous(limits = c(0, 23),
                     breaks = seq(0, 23))                    +
  scale_y_continuous(labels = scales::comma)                 +

  geom_text(aes(x = Hour, y = n, label = comma(n)),
            vjust = -0.5, hjust = 0.5, size = 2, angle = 0) +

  labs(title = "White House Event Start Time Hour - Staff",
       subtitle = paste0(comma(sum(hourCountsLessMissing$n)), " Staff events [",
                         comma(sum(hourCounts$n) -
                               sum(hourCountsLessMissing$n)),
                         " events missing time value] ",
                         "in ", targetMonthName),
       x = "Hour of Day",
       y = "Events",
       caption = c(SOURCE, stamp))                           +
  theme(panel.grid.major.x    = element_blank())
```                                                              

### Tourist

```{r}
hourCounts <-
  events                     |>
  filter(GROUP == "Tourist") |>
  group_by(Hour)             |>
  count()

nrow(hourCounts)
```

```{r}
sum(hourCounts$n)
```

```{r}
hourCountsLessMissing <-
  hourCounts  |>
  filter(Hour != 0)

sum(hourCountsLessMissing$n)
```

```{r TargetEventHourTourist}
ggplot(hourCountsLessMissing)          +
  geom_bar(aes(x = Hour, y = n),
           stat = "identity",
           fill = COLOR_BAR,  color = "black")               +
  
  scale_x_continuous(limits = c(0, 23),
                     breaks = seq(0, 23))                    +
  scale_y_continuous(labels = scales::comma)                 +

  geom_text(aes(x = Hour, y = n, label = comma(n)),
            vjust = -0.5, hjust = 0.5, size = 2, angle = 0) +

  labs(title = "White House Event Start Time Hour - Tourist",
       subtitle = paste0(comma(sum(hourCountsLessMissing$n)), " Tourist events [",
                         comma(sum(hourCounts$n) -
                               sum(hourCountsLessMissing$n)),
                         " events missing time value] ",
                         " in ", targetMonthName),
       x = "Hour of Day",
       y = "Events",
       caption = c(SOURCE, stamp))                           +
  theme(panel.grid.major.x    = element_blank())
```                                                                

### After-Hour Tourists

```{r}
afterHourTouristEvents <- 
  events                     |>
  filter(GROUP == "Tourist",
         Hour  >= 15)        |>
  select(Event)              |>
  unique()

nrow(afterHourTouristEvents)
```

```{r}
afterHourTourists <- 
  wh   |>
  inner_join(afterHourTouristEvents, by = "Event") %>%
  arrange(APPT_START_DATE, VISITEE, VISITOR)

nrow(afterHourTourists)
```

```{r}
if (nrow(afterHourTourists) > 0)
{
  write_xlsx(afterHourTourists,
             paste0("Target-", targetMonth, "-",
                    "White-House-After-Hours-Tourists-",
                    targetMonth, "-",
                    DATA_RELEASE, ".xlsx"))
}
```

## Hour by Month

```{r}
hourCountsByGroupYearMonth <-
  events                                          |>
  mutate(Year  = substr(APPT_START_DATE,  1,  4),
         Month = substr(APPT_START_DATE,  6,  7),
         Hour  = substr(APPT_START_DATE, 12, 13)) |>
  group_by(GROUP, Year, Month, Hour)              |>
  count()                                         |>
  arrange(Year, Month, Hour, GROUP)           

hourCountsByGroupYearMonth |> Show(height = "200px")
```

```{r}
write_xlsx(hourCountsByGroupYearMonth,
           paste0("Target-", targetMonth, "-",
                  "White-House-Event-Start-Hour-by-Group-Month-",
                  DATA_RELEASE, ".xlsx"))
```                                                   

# Visitors by Day

```{r}
byDay <- 
  wh                            |>
  mutate(APPT_DAY = substr(APPT_START_DATE,  1, 10))  |>   # YYYY-MM-DD
  
  group_by(APPT_DAY, GROUP)     |>
  count()                       |>
  ungroup()                     |>
  
  spread(GROUP, n, fill = 0) 
```

```{r}
if (!("FLOTUS" %in% colnames(byDay)))
{
  byDay <- 
    byDay |>
    mutate(FLOTUS = 0) |>
    relocate(FLOTUS, .after = VPOTUS)
}
```


```{r}
byDay <- 
  byDay |>
  mutate( 
          Visitors           = rowSums(across(POTUS:Tourist)),
          PVF                = POTUS + VPOTUS + FLOTUS,
          Day                = mday(APPT_DAY)
        )                    |>
  relocate(Day)

dim(byDay)
```

```{r}
byDay |> Show(bigMark = ",", height = "300px")
```

```{r}
write_xlsx(byDay,
           paste0("Target-", targetMonth, "-",
                  "White-House-Visitors-by-Day-Table-",
                  DATA_RELEASE, ".xlsx"))
```            

```{r}
longFormat <- 
  byDay                   |>
  pivot_longer(
                cols      = c("POTUS", "VPOTUS", "FLOTUS", 
                              "Staff", "Tourist",
                              "Visitors", "PVF"),
                names_to  = "Group",
                values_to = "Visits"
               )

dim(longFormat)
```
Will need to filter for Groups of interest in any plot since there is double counting.

```{r}
longFormat |> head(10) |> Show()
```

## PVL = POTUS + VPOTUS + FLOTUS

Use default blue color for POTUS, but two different sets of qualitative pairs for plots below.

```{r}
POTUS_COLOR <- "#619CFF"
set2 <- brewer.pal( 8, "Set2")  # Brewer qualitative palettes
set3 <- brewer.pal(12, "Set3")
```

```{r VisitorsBidenPotusVPotusFlotus, fig.height = 4, fig.width = 8}
longFormat                                                               |>
  filter(Group %in% c("POTUS", "VPOTUS", "FLOTUS"))                      |>
  mutate(Group = factor(Group, levels = c("FLOTUS", "VPOTUS", "POTUS"))) |>
  ggplot(aes(x = Day, y = Visits, fill = Group))        +
  geom_bar(position = "stack", stat = "identity")       +
  scale_x_continuous(breaks = 1:nrow(eventCountsByDay)) +
  scale_y_continuous(
                      labels = scales::comma
                    )                                   +
                             # FLOTUS    VPOTUS    POTUS
  scale_fill_manual(values = c(set2[5],  set2[6],  POTUS_COLOR))  +

  labs(
        title    = paste0("White House Visitors During Biden Administration in ", targetMonthName),  
        subtitle = paste0("Visitor Counts by Day for POTUS, VPOTUS, FLOTUS "),
        x = "Day of Month", 
        y = "Visitors",
        caption = c(SOURCE, stamp)
      )                                                 +                                   
  theme(panel.grid.major.x    = element_blank(),
        legend.position = "right")
```

```{r VisitorsBidenPotusStaffTourist, fig.height = 4, fig.width = 8}
longFormat                                                             |>
  filter(Group %in% c("PVF", "Staff", "Tourist"))                      |>
  mutate(Group = factor(Group, levels = c("Tourist", "Staff", "PVF"))) |>
  ggplot(aes(x = Day, y = Visits, fill = Group))        +
  geom_bar(position = "stack", stat = "identity")       +
  
  scale_y_continuous(labels = scales::comma)            +
  scale_x_continuous(breaks = 1:nrow(eventCountsByDay)) +
 
   scale_fill_manual(values=c(set3[9],  set3[10], POTUS_COLOR),
                    labels = c("Tourist", "Staff", "POTUS,\nVPOTUS,\nFLOTUS"))  + 
  
  labs(
        title    = paste0("White House Visitors During Biden Administration in ", targetMonthName),  
        subtitle = "Visitor Counts by Day for POTUS/VPOTUS/FLOTUS, Staff, Tourists",
        x = "Day of Month",
        y = "Visitors",
        caption = c(SOURCE, stamp)
      )                                             +                                   
  theme(panel.grid.major.x    = element_blank(),
        legend.position = "right")
```



# Epilog {.tabset .tabset-fade .tabset-pills}

## {.active}

## Session Info

```{r devtoolsSessionInfo}
devtools::session_info()
```

</div>

```{r epilogDeltaTime, echo=FALSE}
time.2 <- Sys.time()
processingTime <- paste("Processing time:", sprintf("%.1f",
                        as.numeric(difftime(time.2,
                                            time.1, units="secs"))), "secs\n")
```

`r processingTime`
`r format(time.2, "%Y-%m-%d %H%M")`
