---
title: "Initial Processing of White House Visitor Logs"
author: "Earl F Glynn<br>"
date: "<small>`r Sys.Date()`</small>"
output:
  html_document:
    code_download: true
    theme: cerulean
    toc: yes
    toc_depth:  3
    toc_float:
      collapsed:  yes
      smooth_scroll: yes
    number_sections: yes
    code_folding:  show
---

Quality control checks, standardization, and descriptive statistics of White House Visitor Log data.

Adapted from script used to process Obama White House visitor records, 2012 - 2016.

```{r setup, echo = FALSE}
# http://biostat.mc.vanderbilt.edu/wiki/Main/KnitrHtmlTemplate
require(Hmisc)    # provides knitrSet and other functions
knitrSet(lang = 'markdown',   # If using blogdown: knitrSet(lang='blogdown')
         fig.align = 'left',
         w = 6.5,
         h = 4.5,
         cache = FALSE)
```

`r hidingTOC(buttonLabel = "Outline")`

```{r startYourEngines, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  comment = NA)

time.1 <- Sys.time()
```

FIXES

* 2023-05-30:  Change VISITEE = "|room" or "rm" (without any first name) to GROUP "Tourist"

# Setup {.tabset .tabset-fade .tabset-pills}

## {.active}

## Constants

```{r}
DATA_RELEASE  <- "2024-06-07"          ##### 1 of 1
BASEDIR       <- "../../../Visitor-Logs-Biden"
```

## Packages

```{r}
library(tidyverse)
library(lubridate)   # parse_date_time, mdy_hm

library(downloader)  # download
library(tools)       # md5sum
```

I/O

```{r IO}
library(readxl)      # read_xlsx
library(readr)       # write_csv
library(writexl)     # write_xlsx
```

Display formatting

```{r Display}
library(kableExtra)  # kable_styling
#library(IRdisplay)   # display_html (needed in Jupyter notebooks)
```

## Helper functions

### Show

```{r Helpers}
Show <- function(data, caption="", bigMark="",
                 height = NULL, width = NULL, ...)
{
  data                                       |>
  kable("html", caption=caption,
        format.args=list(big.mark=bigMark))  |>
  kable_styling(bootstrap_options=c("striped", "bordered", "condensed"),
                position="left",
                full_width=FALSE, ...)       |>
  scroll_box(height = height, width = width)
}
```

### standardizeName

`standardizeName` function removes some of the observed unnecessary variation
in name fields to facilitate comparisons.

In the context of a name an initial is not significant.  Remove all periods
since some name initials have periods and others do not.  Periods are
initially changed to blanks, but extra, leading and trailing spaced are
removed.

Example:

```
standardizeName("           e3.         E..          [(Cumm`ings?           ")
[1] "e e cummings"
```

```{r}
standardizeName <- function(s)
{
  s[s=="n/a"] <- ""

  s <- gsub("/$",  "", s)   # Why do a number of names end with a slash?
  s <- gsub("_|`|0|1|2|3|4|5|6|7|8|9", "", s)  # strange middle names
  s <- gsub("\\-|\\(|\\)|\\[|\\?",     "", s)  # Remove strange characters

  s <- gsub("\\."," ",s)   # Get rid of initials in names
  s <- gsub(" +", " ", s)  # Replace multiple blanks with single blank

  s <- str_trim(s)         # Remove any leading/trailing blanks
  s <- tolower(s)          # force to lower case

  s
}
```

### standarizeLocation

Apply to both MEETING_LOC and MEETING_ROOM fields

```{r}
standardizeLocation <- function(s)
{
  s[s=="n/a"] <- ""

  s <- gsub("/$",  "", s)   # Why do meeting rooms end with a slash?
  s <- gsub(" +", " ", s)  # Replace multiple blanks with single blank

  s <- str_trim(s)         # Remove any leading/trailing blanks
  s <- tolower(s)          # force to lower case

  s
}
```


### ExcelToDateTime

These data formats are from Excel, e.g.,

```
40619.61714 3/17/11 14:48
40621.47378 3/19/11 11:22
```

To convert from Excel to R, express the number of days from the origin to the number of seconds.

```{r}
ExcelToDateTime <- function(excelDates, TZ = "GMT")
{
  as.POSIXct(excelDates * 86400, origin="1899-12-30", tz = TZ)
}
```

# Initial Look

## List of files

```{r}
fileList <- list.files(path       = BASEDIR,
                       pattern    = "^202.*.csv",
                       full.names = TRUE,
                       recursive  = TRUE)
fileList
```

"WAVES" in the filenames above stands for "Workers and Visitors Entry System."

## Read files as strings

Show record counts by file.  These counts include the header record.

```{r}
s <-
  fileList        |>
  map(read_lines)

lengths <- sapply(s, length)
lengths
```

```{r}
sum(lengths)
```

## Verify headers are equivalent

```{r}
headers <- sapply(s, head, 1)
```

Write to .csv file to review in Excel

```{r}
write_lines(paste(basename(fileList), headers, sep = ","),
            "WAVES-ACCESS-RECORDS-Headers.csv")
```

First header

```{r}
headers |> head(1)
```

Most recent header

Starting with the June 2022 file (2022.06_WAVES-ACCESS-RECORDS.csv), the headers were changed to be more "user friendly" but the meaning of the columns did not change.

In the March 2023 file (2023.03_WAVES-ACCESS-RECORDS.csv), headers for last two columns were changed slightly:

* "CALLER_ROOM" -> "Caller Room"

* "RELEASEDATE" -> "Release Date"

There are caseness differences and slight variations in the column names.  Let's use the latest header.

```{r}
headers |> tail(1)
```

## Combine Raw Files

```{r}
s <- c(
        tail(headers, 1),                  # newest header
        sapply(s, tail, -1) |>  unlist()   # remove inconsistent headers
      )

length(s)
```

## Look for duplicate records

Unclear why any records would have exact duplicates.

```{r}
recordCounts <- table(s)

table(recordCounts)
```

Sort and write duplicates to separate files for further study.

```{r}
duplicates <- sort(recordCounts[recordCounts > 1], decreasing=TRUE)

duplicates <- data.frame(counts = as.numeric(duplicates),
                         record = names(duplicates))

dim(duplicates)
```

## Write problems as tab-delimited file since original file has no tabs.

```{r}
PROBLEMS <- "Problems"
if (! file.exists(PROBLEMS))
{
  dir.create(PROBLEMS)
}
```

```{r}
write_tsv(duplicates,
          paste0(PROBLEMS, "/WhiteHouse-Visitors-Duplicate-Records.txt"))
```

```{r, warning = FALSE}
dups <- read_csv(paste0(PROBLEMS, "/WhiteHouse-Visitors-Duplicate-Records.txt"),
                 skip = 1, col_names = FALSE)
```
Look at duplicate counts by release date

```{r}
table( dups |> pull() |> mdy() )
```

## Create unique list of records

```{r}
s <- unique(s)
length(s)
```
## Verify delimiters

All records should have the same number of delimiters.

```{r}
fieldCounts <- count.fields(textConnection(s),
                            sep = ",", comment.char = "", quote = "")
table(fieldCounts)
```

Add headers for the two extra fields, which have been determined to be some sort of `Code` and an `Email` address.

```{r}
s[1] <- paste0(s[1], ",Code,Email")
```

# Read file as fields

Read as character and force any needed conversions

For some reason, textConnection cannot be used here to read from memory, so write to temporary file.

```{r}
writeLines(s, FILENAME_FIXED)
```

Read all fields as character strings.

Carefully control conversions with this "dirty" data.

```{r, warning = FALSE}
d <-
  read_csv(FILENAME_FIXED,
           col_types = cols(.default = "c")) |>
  mutate(
           RELEASEDATE    = mdy(`Release Date`),    # conversions here are OK
           `Total People` = as.numeric(`Total People`)
        )                 |>
  select(-`Release Date`)        # After conversion, keep "old" RELEASEDATE
dim(d)                           # since `Release Date` introduced in March 2023 file
```

```{r}
file.remove(FILENAME_FIXED)
```

## Explore "Extra" fields

Make sure the two extra fields are parsed correctly and not part of RELEASEDATE

RELEASEDATE will be converted to ISO 8601 format later.

```{r}
d                       |>
  group_by(RELEASEDATE) |>
  count()               |>
  Show(bigMark = ",")
```

Why are two records from July 2022 missing `RELEASEDATE`?

```{r}
d |> filter(is.na(RELEASEDATE)) |> Show()
```

### Code

```{r}
table(d$Code, useNA = "ifany")
```

```{r}
table(d$Email,useNA = "ifany")
```

Timeframe for these extra fields.

Problem is limited to July 2022 release date.

```{r}
d                                      |>
  filter(!is.na(Code) & !is.na(Email)) |>
  count(RELEASEDATE)                   |>
  Show()
```

Remove since these fields contain no known useful information

```{r}
d <-
  d  |>
  select(-Code, -Email)
```

## Fields with no information

```{r}
table(d$`Access Type`, useNA = "ifany")
```

```{r}
table(d$POST, useNA = "ifany")
```

```{r}
table(d$`Caller Room`, useNA = "ifany")   # Name changed in March 2023 file
```

Remove since these fields have no useful information

```{r}
d <- d |> select(-`Access Type`, -POST, -`Caller Room`)
dim(d)
```

## Glimpse

```{r}
glimpse(d)
```

The [Wayback Machine from 2012-01-08](https://web.archive.org/web/20120208063341/www.whitehouse.gov/files/disclosures/visitors/WhiteHouse-WAVES-Key-1209.txt) shows some info about the meaning of these fields [with comments]:

```
UIN - Appointment Number [also called Event Number]

BDG NBR – Badge Number

Access Type - Type of access to the complex (VA = Visitor Access)

TOA – Time of Arrival

POA –  Post of Arrival [also called Point of Arrival]

TOD – Time of Departure

POD – Post of Departure [also called Point of Departure]

APPT_MADE_DATE – Date the Appointment was made.

APPT_START_DATE – Date and time for which the appointment was scheduled

APPT_END_DATE – Date and time for which the appointment was scheduled to end

APPT_CANCEL_DATE – Date the appointment was canceled, if applicable

Total_People- The total number of people scheduled for a particular appointment per requestor

LAST_UPDATEDBY – Identifier of officer that updated record

POST – Computer used to enter appointment

LastEntryDate – Most recent update to appointment

TERMINAL_SUFFIX - Identifier of officer that entered appointment

visitee_namelast – Last name of the visitee

visitee_namefirst – First name of the visitee

MEETING_LOC – Building in which meeting was scheduled

MEETING_ROOM – Room in which meeting was scheduled

CALLER_NAME_LAST – Last name of the individual that submitted the WAVES request

CALLER_NAME_FIRST – First name of the individual that submitted the WAVES request

CALLER_ROOM – Room from which the appointment was made

Description – Comments added by the WAVES requestor [no longer part of released records?]
```

The `Description` filed was sometimes used in the Obama records release, but has been dropped.

[121 pages of redacted GW Bush administration "WAVES" visitor records from 2004-2006](https://web.archive.org/web/20121021034409/www.whitehouse.gov/assets/blog/Bush_Admin_Part_2.pdf) can be viewed via the Wayback Machine.  These old records suggest additonal fields are in WAVES but have notw been excluded.

## Cleanup needed for certain fields

Internally use old fields names for variable names modified in July 2022.

This maintains "old" and "new" versions for comparison later if needed.

```{r}
d$NAMELAST  <- standardizeName(d$`Last Name`)
d$NAMEFIRST <- standardizeName(d$`First Name`)
d$NAMEMID   <- standardizeName(d$`Middle Initial`)

d$VISITEE_NAMELAST  <- standardizeName(d$`Visitee Last Name`)
d$VISITEE_NAMEFIRST <- standardizeName(d$`Visitee First Name`)

d$CALLER_NAME_LAST  <- standardizeName(d$`Caller Last Name`)
d$CALLER_NAME_FIRST <- standardizeName(d$`Caller First Name`)
```

## Introduce combined fields

Combined name fields make comparisons easier

```{r}
d$VISITOR <- paste(d$NAMELAST, d$NAMEFIRST, d$NAMEMID, sep="|")
d$VISITEE <- paste(d$VISITEE_NAMELAST, d$VISITEE_NAMEFIRST, sep="|")
d$CALLER  <- paste(d$CALLER_NAME_LAST, d$CALLER_NAME_FIRST, sep="|")
```

Cleanup VISITEE a bit

```{r}
d$VISITEE <- gsub("^NA\\|", "", d$VISITEE)
d$VISITEE <- gsub("\\|NA$", "", d$VISITEE)
```

## Assign `GROUP`

Variation of "Tourist" visitors -- no named person to visit

```
VISITEE,n
office/usss|visitors,5
office|visitor,2
office|visitor's,9
office|visitors,206494
visit,1
visitors,19
visitorsoffice,55
|room,2632
```


```{r}
d <-
  d                |>
  mutate(GROUP =
           case_when(
                      grepl("vpotus|vice|sgotus",  VISITEE)   ~ "VPOTUS",
                      grepl("potus|president",     VISITEE)   ~ "POTUS",
                      grepl("flotus",              VISITEE)   ~ "FLOTUS",

                      grepl("visitor|\\|room|\\|rm", VISITEE) ~ "Tourist",

                      TRUE                                    ~ "Staff"    # more groups?
                    )
        )
```

Kludge Fixes

Not "vice" but staff

```{r}
d$GROUP[d$VISITEE == "secret service|us"] <- "Staff"
```

Likely "whca" is "White House Communications Agency"

```{r}
d$GROUP[d$VISITEE == "service|whca"] <- "Staff"
```

```{r}
d$GROUP[d$VISITEE == "weather|presidential"] <- "Staff"
```

```{r}
table(d$GROUP)
```

### Verify POTUS

President of the United States

```{r}
d |> filter(GROUP == "POTUS") |> count(VISITEE) |> Show()
```

Who visited VPOTUS and Joseph R Biden?

```{r}
d |> filter(VISITEE == "joseph r biden|president") |> Show()
```

### Verify VPOTUS

Vice President of the United States

Note:  `sgotus` is "Second Gentleman of the United States", included with VPOTUS for now

```{r}
d |> filter(GROUP == "VPOTUS") |> count(VISITEE) |> Show()
```

### Verify FLOTUS

First Lady of the United States (Jill Biden)

```{r}
d |> filter(GROUP == "FLOTUS") |> count(VISITEE) |> Show()
```

## Standardize spelling of locations

```{r}
d$MEETING_LOC  <- standardizeLocation(d$`Meeting Location`)
d$MEETING_ROOM <- standardizeLocation(d$`Meeting Room`)
```

LOCATION is combined MEETING_LOC and MEETING ROOM

```{r}
d$LOCATION <- paste(d$MEETING_LOC, d$MEETING_ROOM, sep="|")
```

Add additional standardization (like used with Obama White House data) as needed.

## Fix Dates

Put all dates in standard ISO 8601 format so sorting as character strings results in chronological order.

Historically, dates in White House visitor logs have been a mess and still are.

Use `lubridate` functions now.  Convert in separate steps to isolate problems.

### Appointment Made Date

```{r}
dateLength <- nchar(d$`Appointment Made Date`)
table(dateLength, useNA = "ifany")
```

```{r}
d$`Appointment Made Date`[head(which(dateLength == 13), 3)]
```

```{r}
d$`Appointment Made Date`[head(which(dateLength == 16), 3)]
```

```{r}
d$APPT_MADE_DATE   <- mdy_hm(d$`Appointment Made Date`)
```

### Appointment Start Date

```{r}
dateLength <- nchar(d$`Appointment Start Date`)
table(dateLength, useNA = "ifany")
```

lengths 13-16 work fine for `mdy_hm` function, but short lengths have no time field

Let's look at lengths 8 and 10

```{r}
d$`Appointment Start Date`[head(which(dateLength == 8), 3)]
```

```{r}
d$`Appointment Start Date`[head(which(dateLength == 10), 3)]
```

`truncated = 2` allows conversion to POSIXCt type, but shows no time

```{r}
mdy_hm(d$`Appointment Start Date`[head(which(dateLength == 10), 3)], truncated = 2)
```

```{r}
d$APPT_START_DATE  <- mdy_hm(d$`Appointment Start Date`, truncated = 2)
```

```{r}
tail(d$APPT_START_DATE, 20)
```

### Appointment End Date

```{r}
dateLength <- nchar(d$`Appointment End Date`)
table(dateLength, useNA = "ifany")
```

```{r}
d$APPT_END_DATE    <- mdy_hm(d$`Appointment End Date`)
```

### Appointment Cancel Date

```{r}
dateLength <- nchar(d$`Appointment Cancel Date`)
table(dateLength, useNA = "ifany")
```

What are these unusual lengths, 9 and 11?

```{r}
problemIndices <- which(dateLength %in% c(9, 11))
d$`Appointment Cancel Date`[problemIndices]
```

These are from dates in Excel converted to numbers

```{r}
d                                                                |>
  slice(problemIndices)                                          |>
  select(`Last Name`, `First Name`, starts_with("Appointment"))  |>
  Show()
```

Details of conversion from Excel dates to ISO 8601 strings

```{r}
options(digits = 10)
excelDates <- d$`Appointment Cancel Date`[problemIndices] |> as.numeric()
excelDates
```

```{r}
ExcelToDateTime(excelDates, TZ = "EST5EDT")
```

```{r}
ExcelToDateTime(excelDates, TZ = "EST5EDT") |> str_sub(1, 16)
```

```{r}
d$APPT_CANCEL_DATE <- mdy_hm(d$`Appointment Cancel Date`)
```

```{r}
dateLength <- nchar(d$APPT_CANCEL_DATE)
table(dateLength, useNA = "ifany")
```

Find the three problem dates

```{r}
d$APPT_CANCEL_DATE[problemIndices]
```

```{r}
d$APPT_CANCEL_DATE[problemIndices] <-
  ExcelToDateTime(excelDates, TZ = "EST5EDT") |> str_sub(1, 16)

d$APPT_CANCEL_DATE[problemIndices]
```

```{r}
dateLength <- nchar(d$APPT_CANCEL_DATE)
table(dateLength, useNA = "ifany")
```

```{r}
d$APPT_CANCEL_DATE[which(dateLength == 11)]
```

Why are some now this value?

```{r}
d |> filter(APPT_CANCEL_DATE == as.Date("1900-01-01 UTC")) |> head(2) |> Show()
```

Unconverted `Appointment Cancel Date` orignally was 1/1/1900 0:00.

Let's assume this was some sort of missing value for this field.

```{r}
d$APPT_CANCEL_DATE[which(dateLength == 11)] <- NA
```

### Last Entry Date

```{r}
dateLength <- nchar(d$`Last Entry Date`)
table(dateLength, useNA = "ifany")
```

```{r}
d$`Last Entry Date`[nchar(d$`Last Entry Date`) == 8]  |> head(3)
```

```{r}
d$`Last Entry Date`[nchar(d$`Last Entry Date`) == 9]  |> head(3)
```

```{r}
d$LASTENTRYDATE <- mdy_hm(d$`Last Entry Date`, truncated = 2)
```

### TOA (Time of Arrival)

```{r}
dateLength <- nchar(d$TOA)
table(dateLength, useNA = "ifany")
```

```{r}
d$TOA[which(nchar(d$TOA) ==  8)]  |> head(3)
```

```{r}
d$TOA[which(nchar(d$TOA) == 16)]  |> head(3)
```

```{r}
d$TOA[which(nchar(d$TOA) == 19)]  |> head(3)
```

```{r}
d$TOA[which(nchar(d$TOA) == 20)]  |> head(3)
```
AM/PM messes up normal conversion, so break into two groups

#### No Meridiem

```{r}
NoMeridiem <- grep("AM|PM", d$TOA, invert=TRUE)
length(NoMeridiem)
```

```{r}
d$TOAConverted <- as.POSIXct(rep(NA, nrow(d)))
d$TOAConverted[NoMeridiem] <- mdy_hm(d$TOA[NoMeridiem], truncated = 2)
```

#### Meridiem

```{r}
Meridiem   <- grep("AM|PM", d$TOA, invert=FALSE)
length(Meridiem)
```

Remove commas (only one so far)

```{r}
d$TOA[Meridiem] <- gsub(",", "", d$TOA[Meridiem])
```

Sample conversions

```{r}
d$TOA[Meridiem] |> head(5)
```

Lubridate's `parse_date_time` works better than `strptime` here

```{r}
parse_date_time(d$TOA[Meridiem], "b d y H:Mp", tz = "UTC") |> head(5)
```

```{r}
d$TOAConverted[Meridiem] <-
  parse_date_time(d$TOA[Meridiem], "b d y H:Mp", tz = "UTC")
```

```{r}
with_tz(d$TOAConverted[Meridiem], "UTC") |> head(5)
```

### TOD (Time of Departure)

Mostly missing

```{r}
dateLength <- nchar(d$TOD)
table(dateLength, useNA = "ifany")
```

```{r}
d$TODConverted <- mdy_hm(d$TOD, truncated = 2)
```

## Reorder fields and sort

```{r}
d <-
  d |>
  relocate(APPT_START_DATE, LOCATION, UIN, GROUP,
           VISITEE, VISITOR, `Total People`, CALLER,
           RELEASEDATE, TOA, POA, TOD, POD,
           APPT_MADE_DATE, APPT_END_DATE, APPT_CANCEL_DATE,
           LASTENTRYDATE)                                      |>
  arrange( APPT_START_DATE, LOCATION, UIN, GROUP, VISITEE, VISITOR) |>
  rename(Event = UIN)
```

KLUDGE fix:  `dttm` fields show up like this in a .csv file:  `2021-01-20T12:00:00Z`

But Excel does not handle the `T` and `Z` in these files to allow reformatting.

So, let's force to a character string, which does not have this problem.

Sorting these ISO 8601 date strings still works.

The new strings appear in  a CSV file something like `2021-01-20 12:00:00`

and can be reformated in Excel, e.g., `2021-01-20 12:00`

```{r}
d <-
  d |>
  mutate(
           APPT_START_DATE  = as.character(APPT_START_DATE),
           APPT_MADE_DATE   = as.character(APPT_MADE_DATE),
           APPT_END_DATE    = as.character(APPT_END_DATE),
           APPT_CANCEL_DATE = as.character(APPT_CANCEL_DATE),
           LASTENTRYDATE    = as.character(LASTENTRYDATE),
        )
```

# Write All Info to Single File

```{r}
glimpse(d)
```

Write as csv since may grow over a million (like with Obama administration)

```{r}
write_csv(d,
          paste0("White-House-Visitor-Log-Data-",
                 DATA_RELEASE,
                 ".csv"))
```

# Summary

## Overall

```{r}
overallCounts <-
  d                         |>
  summarize(
              n             = n(),
              nLastName     = n_distinct(NAMELAST),
              nFirstName    = n_distinct(NAMEFIRST),
              nMiddle       = n_distinct(NAMEMID),
              nVISITOR      = n_distinct(VISITOR),

              nVisiteeLast  = n_distinct(VISITEE_NAMELAST),
              nVisiteeFirst = n_distinct(VISITEE_NAMEFIRST),
              nVISITEE      = n_distinct(VISITEE),

              nCallerLast   = n_distinct(CALLER_NAME_LAST),
              nCallerFirst  = n_distinct(CALLER_NAME_FIRST),
              nCALLER       = n_distinct(CALLER),

              nMeetingLoc   = n_distinct(MEETING_LOC),
              nMeetingRoom  = n_distinct(MEETING_ROOM),
              nLOCATION     = n_distinct(LOCATION),

              minApptStart  = min(APPT_START_DATE, na.rm = TRUE),
              maxApptStart  = max(APPT_START_DATE, na.rm = TRUE),

              minPeople     = min(`Total People`, na.rm = TRUE),
              maxPeople     = max(`Total People`, na.rm = TRUE)
           )

overallCounts |>
  Show(bigMark = ",")
```

```{r}
filename <- paste0("White-House-Visitor-Log-Stats-Overall-",
                   DATA_RELEASE,
                   ".xlsx")

write_xlsx(overallCounts, filename)
```


## By Release

```{r}
releaseCounts <-
  d                         |>
  group_by(RELEASEDATE)     |>
  summarize(
              n             = n(),

              nLastName     = n_distinct(NAMELAST),
              nFirstName    = n_distinct(NAMEFIRST),
              nMiddle       = n_distinct(NAMEMID),
              nVISITOR      = n_distinct(VISITOR),

              nVisiteeLast  = n_distinct(VISITEE_NAMELAST),
              nVisiteeFirst = n_distinct(VISITEE_NAMEFIRST),
              nVISITEE      = n_distinct(VISITEE),

              nCallerLast   = n_distinct(CALLER_NAME_LAST),
              nCallerFirst  = n_distinct(CALLER_NAME_FIRST),
              nCALLER       = n_distinct(CALLER),

              nMeetingLoc   = n_distinct(MEETING_LOC),
              nMeetingRoom  = n_distinct(MEETING_ROOM),
              nLOCATION     = n_distinct(LOCATION),

              minApptStart  = min(APPT_START_DATE, na.rm = TRUE),
              maxApptStart  = max(APPT_START_DATE, na.rm = TRUE),

              minPeople     = min(`Total People`, na.rm = TRUE),
              maxPeople     = max(`Total People`, na.rm = TRUE)
           )

releaseCounts |>
  Show(bigMark = ",")
```

```{r}
filename <- paste0("White-House-Visitor-Log-Stats-by-Release-",
                   DATA_RELEASE,
                   ".xlsx")

write_xlsx(releaseCounts, filename)
```

# Frequent

```{r}
table(d$GROUP)
```

## Visitors by Group

### Totals

```{r}
visitorCounts <-
  d                               |>
  select(GROUP, VISITEE, VISITOR) |>
  group_by(VISITOR, GROUP)        |>
  summarize(
              Visits   = n(),
              Visitee  = str_flatten(unique(sort(VISITEE)),
                                     collapse = "; ") |>
                         str_replace_all("\\|", " "),
              .groups = "drop"
           )

nrow(visitorCounts)
```

### By Visit Month

```{r}
visitorCountsByMonth <-
  d                                                 |>
  mutate(Month = substr(APPT_START_DATE, 1, 7))     |>
  select(GROUP, VISITEE, VISITOR, Month)            |>
  group_by(VISITOR, GROUP, Month)                   |>
  summarize( n  = n(), .groups = "drop" )           |>
  spread(Month, n, fill = 0)

nrow(visitorCountsByMonth)
```

```{r}
visitorCounts <-
  visitorCounts |>
  inner_join(visitorCountsByMonth, by = c("VISITOR", "GROUP"))
```

```{r}
visitorCounts %>% head() %>% Show()
```

```{r}
write_csv(visitorCounts,
          paste0("White-House-Visitor-Counts-by-Group-by-Month-",
                 DATA_RELEASE,
                 ".csv"))
```

## Staff Visitee summary of Visitors

POTUS/VPOTUS/FLOTUS visitees can be found in `White-House-Visitor-Log-Data`

```{r}
visiteeCounts <-
  d                               |>
  filter(GROUP == "Staff")        |>
  select(GROUP, VISITEE, VISITOR) |>
  group_by(GROUP, VISITEE)        |>
  summarize(
              Visits   = n(),
              Visitors = str_flatten(unique(sort(VISITOR)),
                                     collapse = "; ") |>
                         str_replace_all("\\|", " "),
              .groups = "drop"
           )           |>
  arrange(GROUP, VISITEE)

nrow(visiteeCounts)
```

Limit strings since Excel restricts to 32767 characters.

For now, only the first 1000 characters

```{r}
visiteeCounts$Visitors[nchar(visiteeCounts$Visitors) > 1000] <-
  paste(substr(visiteeCounts$Visitors[nchar(visiteeCounts$Visitors) > 1000], 1, 1000),
               " . . .")
```

```{r}
visiteeCountsByMonth <-
  d                               |>
  filter(GROUP == "Staff")        |>
  mutate(Month = substr(APPT_START_DATE, 1, 7)) |>
  select(GROUP, VISITEE, VISITOR, Month)        |>
  group_by(GROUP, VISITEE, Month)               |>
  summarize( n  = n(), .groups = "drop" )       |>
  arrange(GROUP, VISITEE)                       |>
  spread(Month, n, fill = 0)

nrow(visiteeCountsByMonth)
```

```{r}
visiteeCounts <-
  visiteeCounts       |>
  inner_join(visiteeCountsByMonth, by = c("GROUP", "VISITEE") )
```

```{r}
visiteeCounts %>% head(3)  %>%  Show()
```

```{r}
visiteeCounts %>% count(Visits) %>% Show(height = "400px")
```

```{r}
write_csv(visiteeCounts,
           paste0("White-House-Staff-Visitee-Counts-by-Month-",
                  DATA_RELEASE,
                  ".csv"))
```

# Descriptive stats for each field

```{r}
COUNTS <- "Counts"  # Directory name for frequency count files
if (! file.exists(COUNTS))
{
  dir.create(COUNTS)
}
```

Save count files as .csv so dates are not converted -- they seem to be if saved in Excel file.  Excel can reformat these dates when loaded.  Use ASCII editor to see actual field values.

Possibly modify these count files for use as filers.

```{r}
fieldSummary <- NULL
for (i in 1:length(colnames(d)))
{
  column <- colnames(d)[i]
  field <-
    d                     |>
    select(x = all_of(i)) |> # kludge using "x" here
    mutate(x = as.character(x))

  counts <-
    field       |>
    group_by(x) |>
    count(x)    |>
    arrange(x)

  colnames(counts)[1] <- column

  filename <- paste(COUNTS, "/", sprintf("%02d",i),"-", column,
                    ".csv",sep="")
  write_csv(counts, filename)
  cat(i, column, "\n")

  zeroLengthField <- (nchar(field$x) == 0)  # trim char strings?
  NMissing <-  sum(zeroLengthField, na.rm = TRUE)

  N <- length(field$x) - NMissing
  NUnique <- length(unique(field$x))

  fieldSummary <-
    rbind(fieldSummary,
          c(i, column,
            as.vector(summary( nchar(field$x[!zeroLengthField]) ) ) |> head(6),
            N, NMissing, NUnique) )  |>
    as_tibble()

}

# wMin .. wMax show field width quantiles
colnames(fieldSummary) <- c("No", "Field", "wMin", "Q1", "wMedian", "wMean",
                            "Q3", "wMax", "N", "Missing", "NUnique")

fieldSummary <-
  fieldSummary   |>
  select(-wMean) |>
  as_tibble()

fieldSummary <-
  fieldSummary |>
  mutate_at(names(fieldSummary)[-2], as.numeric)

dim(fieldSummary)
```

```{r}
write_xlsx(fieldSummary, "WhiteHouse-Visitor-Log-Field-Summary.xlsx")
```

`w` prefix below means field `width`

```{r}
fieldSummary |> Show(height = "400px")
```

What's the deal with missing `VISITEE_NAMELAST`?

Most are for CALLER `ballen|kevin` for Event U93610 on 2022-10-31 with `VISITEE_NAMELAST` missing along with `VISITEE_NAMEFIRST` == "room" and `Meeting LOcation/Meeting Room` of OEOB 1 or `WH SOUTH LAWN`

```{r}
issueVISITEE_NAMELAST <-
  d |>
  filter(nchar(VISITEE_NAMELAST) == 0)

nrow(issueVISITEE_NAMELAST )

write_xlsx(issueVISITEE_NAMELAST,
           paste0(PROBLEMS, "/Missing-VISITEE_NAMELAST.xlsx"))
```


# Epilog {.tabset .tabset-fade .tabset-pills}

## {.active}

## Session Info

```{r devtoolsSessionInfo}
devtools::session_info()
```

</div>

```{r epilogDeltaTime, echo=FALSE}
time.2 <- Sys.time()
processingTime <- paste("Processing time:", sprintf("%.1f",
                        as.numeric(difftime(time.2,
                                            time.1, units="secs"))), "secs\n")
```

`r processingTime`
`r format(time.2, "%Y-%m-%d %H%M")`
