---
title: "Download White House Visitor Logs"
author: "Earl F Glynn<br>"
date: "<small>`r Sys.Date()`</small>"
output:
  html_document:
    theme: cerulean
    toc: yes
    toc_depth:  3
    toc_float:
      collapsed:  yes
      smooth_scroll: yes
    number_sections: yes
    code_folding:  show
---

```{r setup, echo = FALSE}
# http://biostat.mc.vanderbilt.edu/wiki/Main/KnitrHtmlTemplate
require(Hmisc)    # provides knitrSet and other functions
knitrSet(lang = 'markdown',   # If using blogdown: knitrSet(lang='blogdown')
         fig.align = 'left',
         w = 6.5,
         h = 4.5,
         cache = FALSE)
```

`r hidingTOC(buttonLabel = "Outline")`

```{r startYourEngines, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  comment = NA)

time.1 <- Sys.time()
```

[Biden-â Harris Administration Reinstates Visitor Log Policy 2021-05-07](https://www.whitehouse.gov/briefing-room/disclosures/2021/05/07/biden-harris-administration-reinstates-visitor-log-policy-will-be-first-administration-to-post-records-from-first-full-year-in-office/)

* [Voluntary Disclosure Policy for White House Visitor Logs](https://www.whitehouse.gov/voluntary-disclosure/)

* [Visitor Logs](https://www.whitehouse.gov/disclosures/visitor-logs/)

Links to download `.csv` files are in the section `Visitor Logs by Date`

Visitor Log Records Posting Dates:


* [February 2021 Visitor Log Records Posted 2021-06-01](https://www.whitehouse.gov/briefing-room/disclosures/2021/06/01/february-2021-visitor-log-records-posted/)

* [March 2021 Visitor Log Records Posted 2021-06-29](https://www.whitehouse.gov/briefing-room/disclosures/2021/06/29/march-2021-visitor-log-records-posted/)

* [April 2021 Visitor Log Records Posted 2021-07-29](https://www.whitehouse.gov/briefing-room/disclosures/2021/07/29/april-2021-visitor-log-records-posted/)

* [May 2021 Visitor Log Records Posted 2021-08-30](https://www.whitehouse.gov/briefing-room/disclosures/2021/08/30/may-2021-visitor-log-records-posted/)

* [June 2021 Visitor Log Records Posted 2021-10-01](https://www.whitehouse.gov/briefing-room/disclosures/2021/10/01/june-2021-visitor-log-records-posted/)

* [July 2021 Visitor Log Records Posted  2021-11-01](https://www.whitehouse.gov/briefing-room/disclosures/2021/11/01/july-2021-visitor-log-records-posted/)

* [August 2021 Visitor Log Records Posted 2021-12-03 ](https://www.whitehouse.gov/briefing-room/disclosures/2021/12/03/august-2021-visitor-log-records-posted/)

* [September 2021 Visitor Log Records Posted 2022-01-07 ](https://www.whitehouse.gov/briefing-room/disclosures/2022/01/07/september-2021-visitor-log-records-posted/)

* [October 2021 Visitor Log Records Posted 2022-02-04 ](https://www.whitehouse.gov/briefing-room/disclosures/2022/02/04/october-2021-visitor-log-records-posted/)

* [November 2021 Visitor Log Records Posted 2022-03-11 ](https://www.whitehouse.gov/briefing-room/disclosures/2022/03/11/november-2021-visitor-log-records-posted/)

* [December 2021 Visitor Log Records Posted 2022-04-13 ](https://www.whitehouse.gov/briefing-room/disclosures/2022/04/13/december-2021-visitor-log-records-posted/)

# Setup {.tabset .tabset-fade .tabset-pills}

## {.active}

## Constants

```{r}
EXCEL_LIMIT <- 2^20
```

## Packages

```{r}
library(tidyverse)

library(downloader)  # download
library(tools)       # md5sum        
```

Display formatting

```{r Display}
library(kableExtra)  # kable_styling
#library(IRdisplay)   # display_html (needed in Jupyter notebooks)
```         

## Helper functions

```{r Helpers}
Show <- function(data, caption="", bigMark="",
                 height = NULL, width = NULL, ...)
{
  data                                       %>%
  kable("html", caption=caption,
        format.args=list(big.mark=bigMark))  %>%
  kable_styling(bootstrap_options=c("striped", "bordered", "condensed"),
                position="left",
                full_width=FALSE, ...)       %>%
  scroll_box(height = height, width = width) #  %>%
# as.character()                             %>%      # Jupyter Only
# display_html()                                      # Jupyter Only
}
```                           

# Download 2021

```{r StartHere}
FILENAME <- "2021_WAVES-ACCESS-RECORDS.csv"
URL <- paste0("https://www.whitehouse.gov/wp-content/uploads/2022/04/", FILENAME)
download(URL, FILENAME, mode="wb", quiet=TRUE)  
```

```{r}
file.info(FILENAME)              %>%      # metadata about file(s) 
  rownames_to_column("Filename") %>%
  Show(bigMark = ",")
```
```{r}
md5sum(FILENAME)
```

# Epilog {.tabset .tabset-fade .tabset-pills}

## {.active}

## Session Info

```{r devtoolsSessionInfo}
devtools::session_info()
```

</div>

```{r epilogDeltaTime, echo=FALSE}
time.2 <- Sys.time()
processingTime <- paste("Processing time:", sprintf("%.1f",
                        as.numeric(difftime(time.2,
                                            time.1, units="secs"))), "secs\n")
```

`r processingTime`
`r format(time.2, "%Y-%m-%d %H%M")`

