---
title: "White House Visitors - Rework Graphics from Obama Whitehouse Data"
author: "Earl F Glynn<br>"
date: "<small>`r Sys.Date()`</small>"
output:
  html_document:
    code_download: true
    theme: cerulean
    toc: yes
    toc_depth:  3
    toc_float:
      collapsed:  yes
      smooth_scroll: yes
    number_sections: yes
    code_folding:  show     
---

Original analysis: `K:\Data-Farm\US-Government\White-House\Visitor-Logs-Obama\2016\2016-12-30`

Original release:  2016-12-30, `https://www.whitehouse.gov/briefing-room/disclosures/visitor-records`

5.99 millon records released.

AFAIK, visitors for the last three months of Obama's 2nd term  were never released, since there is a three month delay in releases.

Trump is not known to have released Obama's final visitor records.

```{r setup, echo = FALSE}
# http://biostat.mc.vanderbilt.edu/wiki/Main/KnitrHtmlTemplate
require(Hmisc)    # provides knitrSet and other functions
knitrSet(lang = 'markdown',   # If using blogdown: knitrSet(lang='blogdown')
         fig.align = 'left',
         w = 6.5,
         h = 4.5,
         cache = FALSE)
```

`r hidingTOC(buttonLabel = "Outline")`

```{r startYourEngines, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  comment = NA)

time.1 <- Sys.time()
```

Next time:
* Move 

# Setup {.tabset .tabset-fade .tabset-pills}

## {.active}

## Constants

```{r}
SOURCE <-  paste0("Source: www.whitehouse.gov/briefing-room/disclosures/visitor-records\n",
                  "2016-12-30 (Viewable on Wayback Machine)")
stamp   <- paste0("watchdoglab.substack.com\n", format(time.1, "%Y-%m-%d"))
```       


```{r}
EXCEL_LIMIT <- 2^20
```

## Packages

```{r}
library(tidyverse)
library(lubridate)   # parse_date_time, mdy_hm

library(downloader)  # download
library(tools)       # md5sum        
```

I/O

```{r IO}
library(readxl)      # read_xlsx
library(readr)       # write_csv
library(writexl)     # write_xlsx
```

Display formatting

```{r Display}
library(kableExtra)  # kable_styling
#library(IRdisplay)   # display_html (needed in Jupyter notebooks)
```         
Graphics

```{r Graphics}
library(RColorBrewer)
library(scales)      # comma, comma_format
```

## ggplot theme

[ggplot Themes Gallery](https://www.datanovia.com/en/blog/ggplot-themes-gallery/)

```{r ggplotTheme}
theme_set(theme_minimal() +

            theme(
                   axis.text             = element_text(size =  9),
                   axis.title            = element_text(size = 11),

                   plot.caption          = element_text(size =  10,
                                                        hjust = c(0.0,1.0)),
                   plot.caption.position = "plot",  
  
                   plot.title.position   = "plot",
                   plot.title            = element_text(size = 14),
                   plot.subtitle         = element_text(size = 12),
                   
                   strip.background      = element_rect(fill = "aliceblue"),
                   strip.text            = element_text(size = 14),             
            
                   panel.grid.minor.x    = element_blank(),
                   panel.grid.minor.y    = element_blank()   
                 )
         )

COLOR_BAR     <- "skyblue"
COLOR_OUTLINE <- "grey80"
```                                      

## Helper functions

### Show

```{r Helpers}
Show <- function(data, caption="", bigMark="",
                 height = NULL, width = NULL, ...)
{
  data                                       |>
  kable("html", caption=caption,
        format.args=list(big.mark=bigMark))  |>
  kable_styling(bootstrap_options=c("striped", "bordered", "condensed"),
                position="left",
                full_width=FALSE, ...)       |>
  scroll_box(height = height, width = width) 
}
```       

# Obama Visitors by Month

For now, limit to first two years to compare with Biden's first two years.

Create MonthFirst to be used by ggplot functions.

```{r}
byMonthWide <- 
  read_csv("White-House-Visitors-By-Month-Table.csv",
           show_col_types = FALSE)                    |>
  filter(APPT_MONTH < "2011-01")                      |>
  mutate( 
          Visitors = rowSums(across(POTUS:Tourist)),
          Month              = paste0(APPT_MONTH, "-01") |> ymd() 
        )                    |>
  relocate(Month)

dim(byMonthWide)
```

```{r}
byMonthWide |> Show(bigMark = ",", height = "300px")
```

```{r}
longFormat <- 
  byMonthWide             |>
  select(-APPT_MONTH)     |>
  pivot_longer(
                cols      = c("POTUS", "Staff", "Tourist", "Visitors"),
                names_to  = "Group",
                values_to = "Visits"
               )

dim(longFormat)
```
Will need to filter for Groups of interest in any plot since there is double counting.

```{r}
longFormat |> head(10) |> Show()
```

```{r}
POTUS_COLOR <- "#619CFF"
set2 <- brewer.pal( 8, "Set2")  # Brewer qualitative palettes
set3 <- brewer.pal(12, "Set3")
```

```{r ObamaPotusStaffTourist, fig.height = 4, fig.width = 8}
longFormat                                                               |>
  filter(Group %in% c("POTUS", "Staff", "Tourist"))                      |>
  mutate(Group = factor(Group, levels = c("Tourist", "Staff", "POTUS"))) |>
  ggplot(aes(x = Month, y = Visits, fill = Group))  +
  geom_bar(position = "stack", stat = "identity")   +
  
  scale_y_continuous(labels = scales::comma)        +
  scale_x_date(
                date_breaks = "1 month",
                expand = c(0.01, 0), 
                labels = date_format(c("%b",  # terrible kludge here
                                       "%b\n%Y",
                                       rep("%b", 10)))
              )                                     + 
  
  geom_vline(xintercept = c(as.Date("2008-12-17"),
                            as.Date("2009-12-17")),
             color = "black", size = 1,
             linetype = "solid")                    +
  
  annotate("text",
           x = c(as.Date("2009-01-01"),
                 as.Date("2010-01-01")),
           y = Inf,
           label = c("2009", "2010"),
           color = "blue",
           hjust = 0, vjust = 1.5,
           size  = 5)                               +
  
   annotate("text",
           x = as.Date("2009-01-05"),
           y = 40000,
           label = paste0("Obama White House Voluntary Disclosure\n", 
                          "Policy allowed release of visit records\n", 
                          "starting Sept. 15, 2009.  Only requested\n",
                          "records before that date were released."),
           fontface = "italic",
           hjust = 0, vjust = 1,
           size  = 2.5)                               +
  
  scale_fill_manual(values=c(set3[9],  set3[10], POTUS_COLOR),
                    labels = c("Tourist", "Staff", "POTUS,\nVPOTUS,\nFLOTUS"))  + 
  
  labs(
        title    = "White House Visitors During Obama Administration (Jan 2009 - Dec 2010)",
        subtitle = "Visitor Counts by Month for POTUS, Staff, Tourists",
        x = "Month",
        y = "Visitors",
        caption = c(SOURCE, stamp)
      )                                             +                                   
  theme(panel.grid.major.x    = element_blank(),
        legend.position = "right")
```



# Epilog {.tabset .tabset-fade .tabset-pills}

## {.active}

## Session Info

```{r devtoolsSessionInfo}
devtools::session_info()
```

</div>

```{r epilogDeltaTime, echo=FALSE}
time.2 <- Sys.time()
processingTime <- paste("Processing time:", sprintf("%.1f",
                        as.numeric(difftime(time.2,
                                            time.1, units="secs"))), "secs\n")
```

`r processingTime`
`r format(time.2, "%Y-%m-%d %H%M")`
